Functory: A Distributed Computing Library \\ for Objective Caml \thanks{This research was partly supported by the French national project U3CAT (\emph{Unification of Critical C Code Analysis Techniques}, ANR-08-SEGI-021).}

### 2011-03-30 doc/tfp-2011/main.tex
*** empty log message ***

\usepackage{comment,amsmath,graphicx,color,url}

@@

This paper introduces \functory, a generic library for distributed
computing for a widely used functional programming language,
Objective Caml (\Ocaml\ for short). 
This work was initially motivated by the computing needs that
exist in our own research team. Our applications include large-scale
deductive program verification, which amounts to checking the validity
of a large number of logical formulas using a variety of automated
theorem provers~\cite{filliatre07cav}. Our computing infrastructure
consists of a few powerful multi-core machines (typically 8 to 16
cores) and several desktop PCs (typically dual-core). However, for our
application needs, there is no library that helps in exploiting such a
computing infrastructure in our favorite functional programming
language.  Hence we designed and implemented such a library, which is
the subject of this paper.  
% Although this library is implemented in
% \Ocaml, the implementation should be straightforward in any functional
% programming language.

The distributed computing library presented in this paper is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of user-friendly APIs
that allows distributed execution of large-scale parallelizable
computations, very relevant to our application needs (and also
relevant to a variety of real-world applications). Further, the
distributed execution could be over multiple cores in the same machine
or over a network of machines. The most important features of our
library are the following:
\begin{itemize}
\item \emph{Genericity}: 
  it allows various patterns of polymorphic computations;
\item \emph{Simplicity}: switching between multiple cores on the same
  machine and a network of machines is as simple as changing a couple
  of lines of code;
\item \emph{Task distribution and fault-tolerance}: 
  it provides automatic task distribution and
  a robust fault-tolerance mechanisms, thereby
  relieving the user from implementing such routines.
\end{itemize}
The application domain of such a distributed computing library is manyfold. 
% It is worth noting that the library is not targeted 
\functory\ is not targeted at applications running on server farms,
crunching enormous amounts of data, such as Google's
MapReduce~\cite{mapreduce}.  The library is very relevant for
research labs, to make use of available computing resources
efficiently. It serves a variety of users and a wide spectrum of
needs, from desktop PCs to networks of machines.  Typical applications
would involve executing a large number of computation expensive tasks
in a resource-optimal and time-efficient manner. This is also the case
in our research endeavours, that is validating thousands of
verification conditions using automated theorem provers, utilizing the
computing infrastructure to the maximum.

@@

\paragraph{Distributed Computing.}
A typical distributed computing library, as \functory, provides the
following (we borrow some terminology from Google's
MapReduce):
\begin{itemize}
\item A notion of \emph{tasks} which denote atomic computations
  to be performed in a distributed manner; 
\item A set of processes (possibly executing on remote machines)
  called \emph{workers} that perform
  the tasks, producing results;
\item A single process called a \emph{master} which is in charge
  of distributing the tasks among the workers and managing results
  produced by the workers.
\end{itemize}
In addition to the above, distributed computing environments also
implement mechanisms for fault tolerance, efficient storage, and
distribution of tasks. This is required to handle network failures
that may occur, as well as to optimize the usage of machines in the
network. Another concern of importance is the transmission of messages
over the network. This requires efficient 
\emph{marshalling} of data, that is encoding and decoding of data 
for transmission over different computing environments.  It is desirable to
maintain architecture independence while transmitting marshalled data,
as machines in a distributed computing environment often run on
different hardware architectures and make use of different software
platforms. For example, machine word size or endianness may be different
across machines on the network.

\paragraph{A Functional Programming Approach.}
Our work was initially inspired by Google's
MapReduce\footnote{Ironically, Google's approach itself was inspired
  by functional programming primitives.}. However, our functional
programming environment allows us to be more generic. 
The main idea behind our approach is that
workers may implement any polymorphic function:
\begin{ocaml}
  worker: 'a -> 'b
\end{ocaml}
where \of{'a} denotes the type of tasks and \of{'b} the type of results.
Then the master is a
function to handle the results together with
a list of initial tasks:
\begin{ocaml}
  master: ('a -> 'b -> 'a list) -> 'a list -> unit
\end{ocaml}
The function passed to the master is applied whenever a result is
available and may in turn generate new tasks (hence the return type
\of{'a list}).  The master is executed as long as there are pending
tasks.

@@

\JoCaml\ is one of the DFLs which provides communication primitives
(like channels) for facilitating transmission of
computations. However, it does not provide ready-made language
features for fault tolerance, which is indispensable in a distributed
setting. The user has to include code for fault tolerance, as already
demonstrated in some \JoCaml\
library~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
programming language for distributed computing, specialized for web
programming. It provides primitives for transferring control between
the client and the server, as well as low-level primitives for
marshalling the data. As in the case before, ML5 is a programming
language that offers primitives for code mobility, and the code for
distribution of computation and fault-tolerance has to be included by
the user.  ML5 implements type-safe marshalling and
\functory\ does not, though an existing type-safe marshalling library
could be used with \functory. 
% Basically, the task will be transmitted
% along with the corresponding type and on unmarshalling, it could be
% matched for the correct type.
Glasgow Distributed Haskell (GdH)~\cite{GdH} is a pure distributed
functional language that is built on top of Glasgow Haskell and
provides features for distributed computing. It is an extension of
both Glasgow Parallel Haskell, that supports only one process and
multiple threads and Concurrent Haskell that supports multiple
processes. It also offers features for fault-tolerance - error
detection and error recovery primitives in the
language.

CamlP3l~\cite{camlP3l} mixes the features of functional programming
with predefined patterns for parallel computation to offer a parallel
programming environment. Again, it is a programming language offering
primitives for distributing computation to parallel processes and also
to merge the results from parallel executions. Erlang~\cite{erlang} is
a programming language which has features for distribution and fault
tolerance. In particular, it has features for task distribution and is
more well-known for its rich error detection primitives and the
ability to support hot-swapping. The error detection primitives of
Erlang allow nodes to monitor processes in other nodes and also
facilitate automatic migration of tasks in failed nodes to recovered
or active nodes.

@@

There are several implementations of Google's MapReduce in functional
programming languages. But \of{Functory} was just inspired by Google's
MapReduce and is not exactly a MapReduce implementation. The
simplest difference comes from the very fact that \of{Functory} does not
operate on key/value pairs. 
% TODO other difference = absence of data locality
PlasmaMR~\cite{plasma} is an \Ocaml\ implementation of Google's
MapReduce on a distributed file system PlasmaFS. It is able to use
PlasmaFS to its advantage --- the ability of the file system to handle
large files and query functions that implement data locality to
optimize network traffic. However, PlasmaMR does not support fault
tolerance which is indespensible in any distributed computing
application.
% The Disco project~\cite{disco} is an implementation of
% Google's MapReduce in Erlang~\cite{erlang}.
Another MapReduce
implementation in \Ocaml\ is Yohann
Padioleau's~\cite{poor-man-mapreduce}.  It is built on
top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing. Currently, we have less flexibility
w.r.t. deployment of the user program than OCamlMPI; on the other
hand, we provide a more generic API together with fault tolerance.

@@

\begin{comment}
Closest to the approach in this paper is Yohann Padioleau's MapReduce
implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing.  Currently, we have less flexibility
w.r.t. deployment of the user program than OCamlMPI; on the other
hand, we provide a 
more generic API together with fault tolerance.  There are other
distributed computing libraries on top of which one could implement
the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
them. However, \JoCaml\ does not provide fault tolerance, which is
indispensable in a distributed setting. The user has to include code
for fault tolerance, as already demonstrated in some \JoCaml\
experiments~\cite{mandel2008}.

@@

\medskip
The paper is organized as follows. 
Section~\ref{sec:API} describes the API of our library.
Section~\ref{sec:implem} gives implementation details, focusing
on marshaling, network protocol and fault tolerance.
Then Section~\ref{sec:experiments} illustrates the
potential of the presented library through experimental evaluation. 
Finally, Section~\ref{sec:future} outlines some future work.

@@

This section describes our API. 
We start from a simple API which is reduced to a single higher-order
polymorphic function.
Then we explain how this function is actually implemented in terms of
low-level primitives, which are also provided in our API.
The other way round, we also explain how the same function can
be used to implement high-level distribution functions for map and
fold operations. Finally, we explain how our API is
implemented in five different ways, according to five different deployment
scenarios.

@@

The generic distribution function in our API follows the idea
sketched in the introduction. It has the following signature:
\begin{ocaml}
  val compute : 
    worker:('a -> 'b) -> 
    master:('a * 'c -> 'b -> ('a * 'c) list) -> 
    ('a * 'c) list -> unit
\end{ocaml}
Tasks are pairs, of type \of{'a * 'c}, where the first component is
passed to the worker and the second component is local to the master.
The \ocaml{worker} function should be pure\footnote{We mean
  \emph{observationally pure} here but we allow exceptions to be
  raised to signal failures.} 
and is executed in parallel
in all worker processes. The function \ocaml{master}, on the
contrary, can be impure and is only executed sequentially in the master process.
The \ocaml{master} function typically stores results 
in some internal data structure.
Additionally, it may produce new tasks, as a list of type 
\of{('a * 'c) list}, which are then appended to the current set of
pending tasks.

@@

Later, we may want to use a network of machines instead, for example
two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
cores respectively. Similar to the above, we just replace the first
lines with
\begin{ocaml}
  open Network
  let () = declare_workers ~n:4 "orcus"
  let () = declare_workers ~n:8 "belzebuth"
  open Same
\end{ocaml}
Here, \of{Same} is a module which is to be used when master and worker
are running the same executable. It provides a \of{compute} function
which has the same signature as in modules \of{Sequential} and
\of{Cores}, so that the rest of the code remains unchanged. Master and
worker processes are distinguished at run-time using an environment
variable \texttt{WORKER} which is set/unset.

@@

The function \of{compute} above can actually be implemented in terms
of low-level primitives, such as adding a task, adding a worker,
performing one step of communication between master and workers, etc.
These primitives are provided in our API, such that the user can
interact with the execution of the distributed computation.  For
instance, a monitoring-like application can use these primitives to
allow observation and modification of resources (tasks, workers)
during the course of a computation.

A type for distributed computations is introduced:
\begin{ocaml}
  type ('a, 'c) computation
\end{ocaml}
A computation is created with a function \of{create}, which accepts
the same \of{worker} and \of{master} as \of{compute}:
\begin{ocaml}
  val create : 
    worker:('a -> 'b) -> 
    master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a, 'c) computation
\end{ocaml}
Contrary to \of{compute}, it takes no list of tasks and returns
immediately. Tasks can be added later using the following function:
\begin{ocaml}
  val add_task : ('a, 'c) computation -> 'a * 'c -> unit
\end{ocaml}
A function is provided to perform \emph{one step} of a given
computation:
\begin{ocaml}
  val one_step : ('a, 'c) computation -> unit
\end{ocaml}
Calling this function results in one exchange of messages between
master and workers: task assignments to workers, results returned to
the master, etc. A few other functions are provided, such as
\of{status} to query the status of a computation, \of{clear} to remove
all tasks, etc.

@@

The most obvious operation is the traditional map operation over
lists, that is:
\begin{ocaml}
  val map : f:('a -> 'b) -> 'a list -> 'b list
\end{ocaml}
Each task consists of application of function \of{f} to a list element.
More interesting is a combination of map and fold operations.
For instance, we provide different flavors of function
\begin{ocaml}
  val map_fold :
    f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
\end{ocaml}
which, given two functions, an accumulator $a$ and a list $l$, computes
\begin{equation}\label{eq:map-fold}
  \of{fold} ... (\of{fold} (\of{fold} ~ a ~ (\of{f} ~ x_1)) (\of{f} ~ x_2))
  ... (\of{f} ~ x_n)
\end{equation}
for some permutation $[x_1,x_2,...,x_n]$ of the list $l$.
We assume that the \of{f} operations are always performed in parallel.
Regarding \of{fold} operations, we distinguish two cases:
either \of{fold} operations are computationally less expensive
than \of{f} and we perform them locally;
or \of{fold} operations are computationally expensive and we
perform them in parallel.
Thus we provide two functions \of{map_local_fold} and \of{map_remote_fold}.

In the case of \of{map_remote_fold}, only one \of{fold} operation can
be performed at a time (possibly in parallel with \of{f}
operations), as obvious from~(\ref{eq:map-fold}). However, there are
cases where several \of{fold} operations can be performed in parallel,
as early as intermediate results of \of{fold} operations are available.
This is the case when \of{fold} is an associative operation (which
implies that types \of{'b} and \of{'c} are the same).
Whenever
\of{fold} is also commutative, we can perform even more \of{fold}
operations in parallel. Thus our API provides two functions
\of{map_fold_a} and \of{map_fold_ac} for these two particular cases,
with types
\begin{ocaml}
  val map_fold_ac, map_fold_a : 
    f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
\end{ocaml}
This is rather straightforward to derive these five functions from the
generic \of{compute} function; readers interested in details could
refer to the source code.

@@

\subsection{Several APIs in One}

From  the description above, it is clear that our library provides
several APIs of different granularities and different
deployment scenarios. 
Most combinations are meaningful, resulting in
thirteen possible different ways of using our library.
% This is summarized in the following matrix.
% \begin{center}
%   \begin{tabular}{|l|c|c|c|c|c|}
%     \hline
%     & & & \multicolumn{3}{c|}{Network} \\
%     & Seq. & Cores & Same & Poly & Mono \\\hline\hline
%     low-level API & & & & & \\\hline
%     \of{compute}  & & & & & \\\hline
%     high-level API& & & & & \\\hline
%   \end{tabular}
% \end{center}
From the implementation point of view, there is almost no code
duplication. We are using \Ocaml\ functors to derive specific
implementations from generic ones.

@@

We now describe the implementations of the various modules introduced
in Section~\ref{sec:API}. The implementation of the \of{Sequential}
module is straightforward and does not require any explanation.

@@

  \paragraph{\of{Same}.} This module is used when master and workers
  are running the same executable. The master and workers have to be
  differentiated in some manner. We use an environment variable
  \texttt{WORKER} for this purpose. When set, it indicates that the
  executable acts as a worker. At run-time, a worker immediately
  enters a loop waiting for tasks from the master, without even
  getting into the user code.  As explained in Section~\ref{sec:API},
  the master function has the following signature.
  \begin{ocaml}
  val compute : 
    worker:('a -> 'b) -> master:('a * 'c -> 'b -> ('a * 'c) list) -> 
    ('a * 'c) list -> unit
  \end{ocaml}
  The master uses marshalling to send both a closure of type \of{'a ->
    'b} and a task of type \of{'a} to the worker. The resulting
  strings are passed as argument \of{f} and \of{x} in message
  \of{Assign}. Similarly, the worker uses marshalling to send back the
  result of the computation of type \of{'b}, which is the argument
  \of{s} in message \of{Completed}.

  Though the ability to run the same executable helps a lot in
  deploying the program in different machines, it comes at a small
  price. Since the worker is not getting into the user code, closures
  which are transmitted from the master cannot refer to global
  variables in the user code. Indeed, the initialization code for
  these global variables is never reached on the worker side. For
  instance, the Mandelbrot set example could be written as follows:
  \begin{ocaml}
    let max_iterations = 200 
    let worker (xmi, xma, ymi, yma, w, h) = 
      ... draw sub-image using max_iterations ...
  \end{ocaml}
  That is, the global function \of{worker} makes use of the global
  variable \of{max_iterations}. The worker gets the function to
  compute from the master, namely the closure corresponding to
  function \of{worker} in that case, but on the worker side the
  initialization of \of{max_iterations} is never executed.

@@

  \bigskip There are situations where it is not possible to run the
  same executable for master and workers.  For instance, architectures
  or operating systems could be different across the network.  For
  that reason, the \of{Network} module provides two other
  implementations.

  \paragraph{\of{Poly}.} When master and workers are compiled with the
  same version of \Ocaml, we can no longer marshal closures but we can
  still marshal polymorphic values. Indeed, an interesting property of
  marshalling in \Ocaml\ is to be fully architecture-independent, as
  long as a single version of \Ocaml\ is used. It is worth pointing
  out that absence of marshaled closures now enables the use of two
  different programs for master and workers. This is not mandatory,
  though, since master and workers could still be distinguished at
  run-time as in the previous case.

  On the worker side, the main loop is started manually using
  \of{Worker.compute}. The computation to be performed on each task is
  given as an argument to this function. It thus looks as follows:
  \begin{ocaml}
    Worker.compute : ('a -> 'b) -> unit -> unit
  \end{ocaml}
  On the master side, the \of{compute} function is simpler than in the
  previous case, as it has one argument less, and thus has the
  following signature.
  \begin{ocaml}
    Master.compute : 
      master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
  \end{ocaml}
  For realistic applications, where master and workers are completely
  different programs, possibly written by different teams, this is the
  module of choice in our library, since it can still pass polymorphic
  values over the network. The issues of marshalling are automatically
  taken care of by \Ocaml\ run-time.

  The derived API presented in Section~\ref{sec:derived} is adapted to
  deal with the absence of closures. Exactly as the \of{compute}
  function, each API now takes two forms, one for the master and
  another for the workers. For example, \of{map_fold_ac} takes the
  following forms.
  \begin{ocaml}
  Worker.map_fold_ac : f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit 
  Master.map_fold_ac : 'b -> 'a list -> 'b
  \end{ocaml}
  It is the responsibility of the user to ensure type consistency
  between master and workers.

>>>>

\usepackage{comment,amsmath,graphicx,color,url,hyperref}

@@

This paper introduces \functory, a generic library for distributed
computing for a widely used functional programming language,
Objective Caml (\Ocaml\ for short). 
This work was initially motivated by the computing needs that
exist in our own research team. Our applications include large-scale
deductive program verification, which amounts to checking the validity
of a large number of logical formulas using a variety of automated
theorem provers~\cite{filliatre07cav}. Our computing infrastructure
consists of a few powerful multi-core machines (typically 8 to 16
cores) and several desktop PCs (typically dual-core). However, for our
application needs, no existing library provides a polymorphic API with
usual map/fold higher-order operations, built-in fault-tolerance, and
the ability to easily switch between multi-core and network
infrastructures.
Hence we designed and implemented such a library, which is
the subject of this paper.  
The library is available at 
\href{http://functory.lri.fr/}{\url{http://functory.lri.fr/}}.

The distributed computing library presented in this paper is not a
library that helps in parallelizing computations. Rather, it provides
facilities for reliable, distributed execution of parallelizable
computations. In particular, it provides a set of user-friendly APIs
that allows distributed execution of large-scale parallelizable
computations, very relevant to our application needs (and also
relevant to a variety of real-world applications). Further, the
distributed execution could be over multiple cores in the same machine
or over a network of machines. The most important features of our
library are the following:
\begin{itemize}
\item \emph{Genericity}: 
  it allows various patterns of polymorphic computations;
\item \emph{Simplicity}: switching between multiple cores on the same
  machine and a network of machines is as simple as changing a couple
  of lines of code;
\item \emph{Task distribution and fault-tolerance}: 
  it provides automatic task distribution and
  a robust fault-tolerance mechanisms, thereby
  relieving the user from implementing such routines.
\end{itemize}
The application domain of such a distributed computing library is manyfold. 
It serves a variety of users and a wide spectrum of
needs, from desktop PCs to networks of machines.  Typical applications
would involve executing a large number of computation expensive tasks
in a resource-optimal and time-efficient manner. This is also the case
in our research endeavours, that is validating thousands of
verification conditions using automated theorem provers, utilizing the
computing infrastructure to the maximum.
It is worth noting that 
\functory\ is not targeted at applications running on server farms,
crunching enormous amounts of data, such as Google's
MapReduce~\cite{mapreduce}.

@@

\paragraph{Distributed Computing.}
A typical distributed computing library, as \functory, provides the
following (we borrow some terminology from Google's
MapReduce):
\begin{itemize}
\item A notion of \emph{tasks} which denote atomic computations
  to be performed in a distributed manner; 
\item A set of processes (possibly executing on remote machines)
  called \emph{workers} that perform
  the tasks, producing results;
\item A single process called a \emph{master} which is in charge
  of distributing the tasks among the workers and managing results
  produced by the workers.
\end{itemize}
In addition to the above, distributed computing environments also
implement mechanisms for fault-tolerance, efficient storage, and
distribution of tasks. This is required to handle network failures
that may occur, as well as to optimize the usage of machines in the
network. Another concern of importance is the transmission of messages
over the network. This requires efficient 
\emph{marshalling} of data, that is encoding and decoding of data 
for transmission over different computing environments.  It is desirable to
maintain architecture independence while transmitting marshalled data,
as machines in a distributed computing environment often run on
different hardware architectures and make use of different software
platforms. For example, machine word size or endianness may be different
across machines on the network.

\paragraph{A Functional Programming Approach.}
Our work was initially inspired by Google's
MapReduce\footnote{Ironically, Google's approach itself was inspired
  by functional programming primitives.}. However, our functional
programming environment allows us to be more generic. 
The main idea behind our approach is that
workers may implement any polymorphic function:
\begin{ocaml}
  worker: 'a -> 'b
\end{ocaml}
where \of{'a} denotes the type of tasks and \of{'b} the type of results.
Then the master is a
function to handle the results together with
a list of initial tasks:
\begin{ocaml}
  master: ('a -> 'b -> 'a list) -> 'a list -> unit
\end{ocaml}
The function passed to the master is applied whenever a result is
available. The first argument is the task (of type \of{'a}) and the
second one its result (of type \of{'b}). 
It may in turn generate new tasks, hence the return type
\of{'a list}.  
The master is executed as long as there are pending tasks.

@@

\JoCaml\ is one of the DFLs which provides communication primitives
(like channels) for facilitating transmission of
computations. However, it does not provide ready-made language
features for fault-tolerance, which is indispensable in a distributed
setting. The user has to include code for fault-tolerance, as already
demonstrated in some \JoCaml\
library~\cite{mandel2008}. ML5~\cite{ML5}, a variant of ML, is a
programming language for distributed computing, specialized for web
programming. It provides primitives for transferring control between
the client and the server, as well as low-level primitives for
marshalling the data. As in the case before, ML5 is a programming
language that offers primitives for code mobility, and the code for
distribution of computation and fault-tolerance has to be included by
the user.  ML5 implements type-safe marshalling and
\functory\ does not, though an existing type-safe marshalling library
could be used with \functory. 
% Basically, the task will be transmitted
% along with the corresponding type and on unmarshalling, it could be
% matched for the correct type.
Glasgow Distributed Haskell (GdH)~\cite{GdH} is a pure distributed
functional language that is built on top of Glasgow Haskell and
provides features for distributed computing. It is an extension of
both Glasgow Parallel Haskell, that supports only one process and
multiple threads and Concurrent Haskell that supports multiple
processes. It also offers features for fault-tolerance - error
detection and error recovery primitives in the
language.

CamlP3l~\cite{camlP3l} mixes the features of functional programming
with predefined patterns for parallel computation to offer a parallel
programming environment. Again, it is a programming language offering
primitives for distributing computation to parallel processes and also
to merge the results from parallel executions. Erlang~\cite{erlang} is
a programming language which has features for distribution and 
fault-tolerance. In particular, it has features for task distribution and is
more well-known for its rich error detection primitives and the
ability to support hot-swapping. The error detection primitives of
Erlang allow nodes to monitor processes in other nodes and also
facilitate automatic migration of tasks in failed nodes to recovered
or active nodes.

@@

There are several implementations of Google's MapReduce in functional
programming languages. But \of{Functory} was just inspired by Google's
MapReduce and is not exactly a MapReduce implementation. The
simplest difference comes from the very fact that \of{Functory} does not
operate on key/value pairs. 
% TODO other difference = absence of data locality
PlasmaMR~\cite{plasma} is an \Ocaml\ implementation of Google's
MapReduce on a distributed file system PlasmaFS. It is able to use
PlasmaFS to its advantage --- the ability of the file system to handle
large files and query functions that implement data locality to
optimize network traffic. However, PlasmaMR does not support 
fault-tolerance which is indespensible in any distributed computing
application.
% The Disco project~\cite{disco} is an implementation of
% Google's MapReduce in Erlang~\cite{erlang}.
Another MapReduce
implementation in \Ocaml\ is Yohann
Padioleau's~\cite{poor-man-mapreduce}.  It is built on
top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing. Currently, we have less flexibility
w.r.t. deployment of the user program than OCamlMPI; on the other
hand, we provide a more generic API together with fault-tolerance.

@@

\begin{comment}
Closest to the approach in this paper is Yohann Padioleau's MapReduce
implementation in \Ocaml~\cite{poor-man-mapreduce}.  It is built on
top of OCamlMPI~\cite{ocamlMPI}, while our approach uses a homemade
protocol for message passing.  Currently, we have less flexibility
w.r.t. deployment of the user program than OCamlMPI; on the other
hand, we provide a 
more generic API together with fault-tolerance.  There are other
distributed computing libraries on top of which one could implement
the library discussed in this paper. \JoCaml~\cite{jocaml} is one of
them. However, \JoCaml\ does not provide fault-tolerance, which is
indispensable in a distributed setting. The user has to include code
for fault-tolerance, as already demonstrated in some \JoCaml\
experiments~\cite{mandel2008}.

@@

\medskip
The paper is organized as follows. 
Section~\ref{sec:API} describes the API of our library.
Section~\ref{sec:implem} gives implementation details, focusing
on marshaling, network protocol and fault-tolerance.
Then Section~\ref{sec:experiments} illustrates the
potential of the presented library through experimental evaluation. 
Finally, Section~\ref{sec:future} outlines some future work.

@@

This section describes our API. 
We start from a simple API which is reduced to a single higher-order
polymorphic function.
Then we explain how this function is actually implemented in terms of
low-level primitives, which are also provided in our API.
Conversely, we also explain how the same function can
be used to implement high-level distribution functions for map and
fold operations. Finally, we explain how our API is
implemented in five different ways, according to five different deployment
scenarios.

@@

The generic distribution function in our API follows the idea
sketched in the introduction. It has the following signature:
\begin{ocaml}
  val compute: 
    worker:('a -> 'b) -> 
    master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
\end{ocaml}
Tasks are pairs, of type \of{'a * 'c}, where the first component is
passed to the worker and the second component is local to the master.
The \ocaml{worker} function should be pure\footnote{We mean
  \emph{observationally pure} here but we allow exceptions to be
  raised to signal failures.} 
and is executed in parallel
in all worker processes. The function \ocaml{master}, on the
contrary, can be impure and is only executed sequentially in the master process.
The \ocaml{master} function typically stores results 
in some internal data structure.
Additionally, it may produce new tasks, as a list of type 
\of{('a * 'c) list}, which are then appended to the current set of
pending tasks.

@@

Later, we may want to use a network of machines instead, for example
two machines called \texttt{orcus} and \texttt{belzebuth} with 4 and 8
cores respectively. Similar to the above, we just replace the first
lines with
\begin{ocaml}
  open Network
  let () = declare_workers ~n:4 "orcus"
  let () = declare_workers ~n:8 "belzebuth"
  open Same
\end{ocaml}
Here, \of{Same} is a module which is to be used when master and worker
are running the same executable. It provides a \of{compute} function
which has the same signature as in modules \of{Sequential} and
\of{Cores}, so that the rest of the code remains unchanged. Master and
worker processes are distinguished at runtime using an environment
variable \texttt{WORKER} which is set/unset.

@@

The function \of{compute} above can actually be implemented in terms
of low-level primitives, such as adding a task, adding a worker,
performing some communication between master and workers, etc.
These primitives are provided in our API, such that the user can
interact with the execution of the distributed computation.  For
instance, a monitoring-like application can use these primitives to
allow observation and modification of resources (tasks, workers)
during the course of a computation.
A type for distributed computations is introduced:
\begin{ocaml}
  type ('a, 'c) computation
\end{ocaml}
A computation is created with a function \of{create}, which accepts
the same \of{worker} and \of{master} as \of{compute}:
\begin{ocaml}
  val create: worker:('a -> 'b) -> 
    master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a, 'c) computation
\end{ocaml}
Contrary to \of{compute}, it takes no list of tasks and returns
immediately. Tasks can be added later using the following function:
\begin{ocaml}
  val add_task: ('a, 'c) computation -> 'a * 'c -> unit
\end{ocaml}
A function is provided to perform \emph{one step} of a given
computation:
\begin{ocaml}
  val one_step: ('a, 'c) computation -> unit
\end{ocaml}
Calling this function results in one exchange of messages between
master and workers: task assignments to workers, results returned to
the master, etc. A few other functions are provided, such as
\of{status} to query the status of a computation, \of{clear} to remove
all tasks, etc.

@@

The most obvious operation is the traditional map operation over
lists, that is \of{val map: f:('a -> 'b) -> 'a list -> 'b list}.
Each task consists of the application of function \of{f} to a list element.
More interesting is a combination of map and fold operations.
For instance, we provide different flavors of function
\begin{ocaml}
  val map_fold: f:('a -> 'b) -> fold:('c -> 'b -> 'c) -> 'c -> 'a list -> 'c
\end{ocaml}
which, given two functions, an accumulator $a$ and a list $l$, computes
\begin{equation}\label{eq:map-fold}
  \of{fold} ... (\of{fold} (\of{fold} ~ a ~ (\of{f} ~ x_1)) (\of{f} ~ x_2))
  ... (\of{f} ~ x_n)
\end{equation}
for some permutation $[x_1,x_2,...,x_n]$ of the list $l$.
We assume that the \of{f} operations are always performed in parallel.
Regarding \of{fold} operations, we distinguish two cases:
either \of{fold} operations are computationally less expensive
than \of{f} and we perform them locally;
or \of{fold} operations are computationally expensive and we
perform them in parallel.
Thus we provide two functions \of{map_local_fold} and \of{map_remote_fold}.

In the case of \of{map_remote_fold}, only one \of{fold} operation can
be performed at a time (possibly in parallel with \of{f}
operations), as obvious from~(\ref{eq:map-fold}). However, there are
cases where several \of{fold} operations can be performed in parallel,
as early as intermediate results of \of{fold} operations are available.
This is the case when \of{fold} is an associative operation (which
implies that types \of{'b} and \of{'c} are the same).
Whenever
\of{fold} is also commutative, we can perform even more \of{fold}
operations in parallel. Thus our API provides two functions
\of{map_fold_a} and \of{map_fold_ac} for these two particular cases,
with types
\begin{ocaml}
  val map_fold_ac, map_fold_a: 
    f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> 'b -> 'a list -> 'b 
\end{ocaml}
It is rather straightforward to derive these five functions from the
generic \of{compute} function; readers interested in details could
refer to the source code.

@@

\subsection{Several Libraries in One}

From  the description above, it is clear that our library provides
several APIs of different granularities, as well as several
implementations for various deployment scenarios. 
Most combinations are meaningful, resulting in
thirteen possible different ways of using our library.
For instance, one may use the low-level API on a single multi-core
machine, or use the high-level API on a network of machines all
running the same executable, etc.
From the implementation point of view, there is almost no code
duplication. We are using \Ocaml\ functors to derive specific
implementations from generic ones.

@@

The implementation of the \of{Sequential}
module is straightforward and does not require any explanation.
The \of{Cores} module is implemented with \unix\ processes, using
the \of{fork} and \of{wait} system calls provided by the \of{Unix}
library of \Ocaml. We do not describe this implementation but
rather focus on the more interesting module \of{Network}.

@@

  \paragraph{\of{Same}.} This module is used when master and workers
  are running the same executable. The master and workers have to be
  differentiated in some manner. We use an environment variable
  \texttt{WORKER} for this purpose. When set, it indicates that the
  executable acts as a worker. At runtime, a worker immediately
  enters a loop waiting for tasks from the master, without even
  getting into the user code.  As explained in Section~\ref{sec:API},
  the master function has the following signature.
  \begin{ocaml}
  val compute: worker:('a -> 'b) -> 
    master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
  \end{ocaml}
  The master uses marshalling to send both a closure of type \of{'a ->
    'b} and a task of type \of{'a} to the worker. The resulting
  strings are passed as argument \of{f} and \of{x} in message
  \of{Assign}. Similarly, the worker uses marshalling to send back the
  result of the computation of type \of{'b}, which is the argument
  \of{s} in message \of{Completed}.

  Though the ability to run the same executable helps a lot in
  deploying the program in different machines, it comes at a small
  price. Since the worker is not getting into the user code, closures
  which are transmitted from the master cannot refer to global
  variables in the user code. Indeed, the initialization code for
  these global variables is never reached on the worker side. For
  instance, some code for drawing Mandelbrot's set
  could be written as follows:
  \begin{ocaml}
    let max_iterations = 200 
    let worker si = ... draw sub-image si using max_iterations ...
  \end{ocaml}
  That is, the global function \of{worker} makes use of the global
  variable \of{max_iterations}. The worker gets the function to
  compute from the master, namely the closure corresponding to
  function \of{worker} in that case, but on the worker side the
  initialization of \of{max_iterations} is never executed.

@@

  There are situations where it is not possible to run the
  same executable for master and workers.  For instance, architectures
  or operating systems could be different across the network.  For
  that reason, the \of{Network} module provides two other
  implementations.

  \paragraph{\of{Poly}.} When master and workers are compiled with the
  same version of \Ocaml, we can no longer marshal closures but we can
  still marshal polymorphic values. Indeed, an interesting property of
  marshalling in \Ocaml\ is to be fully architecture-independent, as
  long as a single version of \Ocaml\ is used. It is worth pointing
  out that absence of marshaled closures now enables the use of two
  different programs for master and workers. This is not mandatory,
  though, since master and workers could still be distinguished at
  runtime as in the previous case.

  On the worker side, the main loop is started manually using
  \of{Worker.compute}. The computation to be performed on each task is
  given as an argument to this function. It thus looks as follows:
  \begin{ocaml}
    Worker.compute: ('a -> 'b) -> unit -> unit
  \end{ocaml}
  On the master side, the \of{compute} function is simpler than in the
  previous case, as it has one argument less, and thus has the
  following signature.
  \begin{ocaml}
    Master.compute: 
      master:('a * 'c -> 'b -> ('a * 'c) list) -> ('a * 'c) list -> unit
  \end{ocaml}
  For realistic applications, where master and workers are completely
  different programs, possibly written by different teams, this is the
  module of choice in our library, since it can still pass polymorphic
  values over the network. The issues of marshalling are automatically
  taken care of by the \Ocaml\ runtime.

  The derived API presented in Section~\ref{sec:derived} is adapted to
  deal with the absence of closures. Exactly as the \of{compute}
  function, each API now takes two forms, one for the master and
  another for the workers. For example, \of{map_fold_ac} takes the
  following forms.
  \begin{ocaml}
  Worker.map_fold_ac: f:('a -> 'b) -> fold:('b -> 'b -> 'b) -> unit 
  Master.map_fold_ac: 'b -> 'a list -> 'b
  \end{ocaml}
  It is the responsibility of the user to ensure consistency
  between master and workers.
